Perturbing SU($n$)-symmetric interactions

### 2020-02-06 collective_sun/sun_perturbation.tex
final changes to diagram definitions

To simplify the product in \eqref{eq:sym_prod_proj}, we first write
\begin{align}
  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
  = \prod_{j\in\ZZ_p} \f1{M_j!} \sum_{k\in\ZZ_N^{M_j}}
  w_j\p{k} O_j\p{k},
  \label{eq:sym_prod_vec}
\end{align}
where we extend the definitions of $w_j\p{k}$ and $O_j\p{k}$ to all
vectors $k\in\ZZ_N^{M_j}$ by defining symmetric dimension-$M_j$
tensors $w_j$ and $O_j$ (i.e.~tensors with $M_j$ indices, with $O_j$
an {\it operator-valued} tensor) for which $w_j\p{k}=O_j\p{k}=0$ if
$k$ contains any repeated symbols.  We then collect terms whose
operator content is equivalent up to a permutation of spins.  To this
end, we classify terms in \eqref{eq:sym_prod_vec} by the numbers $g_S$
of indices shared by all tensors $w_j,O_j$ with $j\in S\subset\ZZ_p$.
For example, a term of the form
$w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}$ with distinct indices
$\p{a,b,c,d,e}\in\C_N\p{5}$ would have
\begin{align}
  g_{\set{1}} = 1,
  &&
  g_{\set{1,2,3}} = 1,
  &&
  g_{\set{1,3}} = 1,
  &&
  g_{\set{2,3}} = 2,
\end{align}
and $g_S=0$ for all other subsets $S\subset\ZZ_3$.  This index
assignment can be represented by the Venn diagram
\begin{align}
  \diagram{example_triple},
  \label{eq:venn_diagram}
\end{align}
where $g_S$ is determined the number of dots at the intersection of
circles $j\in S$.  We denote an assignment of distinct indices
according to choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ by $g$, where
$\PP\p{\ZZ_p}$ is the power set (i.e.~set of all subsets) of $\ZZ_p$.
Given a vector $\m M$ of the dimensions $M_j$, we then denote the set
of all valid index assignments $g$ by $\G\p{\m M}$.  For consistency,
all valid index assignments $g\in\G\p{\m M}$ must assign exactly $M_j$
indices to the dimension-$M_j$ tensor $w_j$, i.e.~
\begin{align}
  \sum_{S\in\PP\p{\ZZ_p}\,:\,j\in S} g_S
  = \sum_{R\in\PP\p{\ZZ_p\setminus\set{j}}} g_{\set{j}\cup R}
  = M_j
\end{align}
for all $j\in\ZZ_p$.  In order to make the index assignments
determined by choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ unique, we
also enforce that $g_{\set{}}=0$ for all $g\in\G\p{\m M}$, which is
equivalent to saying that every dot in the corresponding Venn diagram
must lie within some circle.  The set of valid index assignments
$\G\p{\m M}$ is then essentially the set of all Venn diagrams of the
form in \eqref{eq:venn_diagram} with $p=\dim\p{\m M}$ circles and a
total of $M_j$ dots within circle $j$.

@@

The simplest diagrams have $g_S$ dots (``$\bullet$'') at the
intersection of the circles $j\in S$, and are proportional to the
associated scalar $w\p{g}$ defined in \eqref{eq:diagram_factors}, e.g.
\begin{align}
  \diagram{example_triple}
  \equiv \sp{\prod_{S\in\PP\p{\ZZ_p}} g_S!} w\p{g}
  = \sum_{\substack{a,b,c,d,e\in\ZZ_N\\
      \abs{\set{a,b,c,d,e}}=5}}
  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}.
  \label{eq:diagram_definition}
\end{align}
The proportionality factor accounts for the number of ways to permute
the indices within each region $S\in\PP\p{\ZZ_p}$.  Including this
symmetry factor in the definition of a diagram is a matter of choice
that will ultimately make it easier to work with diagrams.  A diagram
such at that in \eqref{eq:diagram_definition} with $\abs{g}$ dots
nominally takes $\O\p{N^{\abs{g}}}$ time to compute.  In practice,
this computational cost can be prohibitive for performing numerical
simulations of systems of any appreciable size.  If possible, we
therefore wish to simplify diagrams in order to reduce their
computational complexity.  To this end, we first need to define simple
modifications of diagrams in which dots may be replaced by empty dots
(``$\circ$'') or crosses (``$\bm\times$'').

In any diagram (with filled dots, empty dots, or crosses), the values
of indices associated with filled dots are always summed as if the
diagram had only dots, without regard for all other indices.  Empty
dots in a diagram then correspond to indices that are summed over all
of $\ZZ_N$ independently and without additional constraints, e.g.
\begin{align}
  \diagram{example_triple_oo_2} \equiv
  \sum_{a,c\in\ZZ_N}
  \sum_{\substack{b,d,e\in\ZZ_N\\\abs{\set{b,d,e}}=3}}
  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}.
\end{align}
Crosses, meanwhile, correspond to indices that are {\it maximally
  dependent} on filled dots: ``cross indices'' are summed {\it only}
over the values of ``filled dot indices'', e.g.
\begin{align}
  \diagram{example_triple_xx_1}
  \equiv \sum_{\substack{a,d,e\in\ZZ_N\\\abs{\set{a,d,e}}=3}}
  \sum_{b,c\in\set{a,d,e}}
  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}.
\end{align}
Empty dots and crosses allow us to decompose diagrams into sums of
different diagrams that (may) require less time to evaluate.  We will
essentially use two tricks to decompose diagrams: eliminating filled
dots, and eliminating crosses.

Given a diagram containing only dots (filled or empty), we can
decompose any filled dot into an empty dot and a cross.  For example
\begin{align}
  \diagram{example_triple}
  = \diagram{example_triple_o_top} - \diagram{example_triple_x_top},
  \label{eq:example_decomposition}
\end{align}
where
\begin{align}
  \diagram{example_triple}
  \equiv \sum_{\substack{a,b,c,d,e\in\ZZ_N\\\abs{\set{a,b,c,d,e}}=5}}
  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e},
\end{align}
\begin{align}
  \diagram{example_triple_x_top}
  \equiv \sum_{\substack{b,c,d,e\in\ZZ_N\\\abs{\set{b,c,d,e}}=4}}
  \sum_{a\in\set{b,c,d,e}}
  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e},
\end{align}
\begin{align}
  \diagram{example_triple_o_top}
  \equiv \sum_{\substack{b,c,d,e\in\ZZ_N\\\abs{\set{b,c,d,e}}=4}}
  w_1\p{\circ,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e},
  \label{eq:example_triple_x_top}
\end{align}
with
\begin{align}
  w_1\p{\circ,b,c} \equiv \sum_{a\in\ZZ_N} w_1\p{a,b,c},
\end{align}
The decomposition in \eqref{eq:example_decomposition} essentially
breaks up a sum over $a\in\ZZ_N\setminus\set{b,c,d,e}$ into a
difference of sums over $a\in\ZZ_N$ and $a\in\set{b,c,d,e}$.  The sum
over $a\in\ZZ_N$ to compute $w_1\p{\circ,b,c}$ has $\O\p{N^3}$ cost,
and can be performed prior to computing the diagram in
\eqref{eq:example_triple_x_top}.  The decomposition in
\eqref{eq:example_decomposition} thereby reduces the calculation of
one $\O\p{N^5}$ diagram to the calculation of two $\O\p{N^4}$
diagrams.  A similar reduction by a factor of $\sim N$ can be achieved
by decomposing any dot in a region $S\in\PP\p{\ZZ_p}$ addressing only
one tensor, i.e.~with $\abs{S}=1$.

While it is possible, in principle, to derive rules for decomposing
filled dots in a diagram that contains crosses, it is much easier to
simply eliminate crosses from a diagram altogether.  To eliminate a
cross from a diagram, we first note that the corresponding index need
not sum over the values of indices in any tensor that contains that
index; that is (pay attention to the second sum in each line),
\begin{align}
  \diagram{example_triple_x_top_elim}
  &\equiv \sum_{\substack{b,c,d,e,f\in\ZZ_N\\\abs{\set{b,c,d,e,f}}=5}}
  \sum_{a\in\set{b,c,d,e,f}}
  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,f} \\
  &= \sum_{\substack{b,c,d,e,f\in\ZZ_N\\\abs{\set{b,c,d,e,f}}=5}}
  \sum_{a\in\set{d,e,f}}
  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,f}.
  \label{eq:x_elim}
\end{align}
For each remaining index addressed by a cross, meanwhile, the cross
simply forces the corresponding indices to be equal, so
\begin{align}
  \diagram{example_triple_x_top_elim}
  = 2 \diagram{example_triple_x_lft_elim}
  + \diagram{example_triple_x_rht_elim},
\end{align}
where the factor of 2 appears simply to account for both of
$a\in\set{d,e}$ in \eqref{eq:x_elim}.  The rules derived here for
replacing filled dots and crosses will be used in the following
sections to simplify diagrams and minimize the computational cost of
their evaluation.

@@

Calculating the second-order effective Hamiltonian $H_2^{(2)}$ induced
on the permutationally symmetric manifold $\M_0$ by the perturbation
$\V_2$ in \eqref{eq:perturbations} requires us to simplify a product
of the form
\begin{multline}
  \P_0 \sp{\sum_{k<\ell} \sum_{p<q}
    v_{k\ell} w_{pq} O_{k\ell} Q_{pq}} \P_0 \\
  \EQFS \f14 \diagram{two_body_0} O_{1,2} Q_{3,4}
  + \diagram{two_body_1} O_{1,2} Q_{1,3}
  + \f12 \diagram{two_body_2} O_{1,2} Q_{1,2},
  \label{eq:POQP_start}
\end{multline}
where $k,\ell,p,q$ index individual spins; $O$ and $Q$ are
permutationally-symmetric two-spin operators; the coefficients
$v_{k\ell},w_{pq}$ satisfy \eqref{eq:cond_2} with eigenvalue
$\Delta\ne0$, which implies that they are mean-zero; and the factors
of $1/4$ and $1/2$ are included to cancel out the symmetry factors
that appear in the definition of a diagram.  Defining
\begin{align}
  \v{\m m} \equiv \sum_{p<q} m_{pq} \ket{\set{p,q}},
  &&
  \v m \equiv \sum_p m_p \ket{p},
  &&
  m_p \equiv \sum_q m_{pq},
\end{align}
for $m\in\set{v,w}$, we can write
\begin{align}
  \diagram{two_body_2} = \sum_{p,q\in\ZZ_N} v_{pq} w_{pq}
  = 2 \sum_{p,q\in\C_N\p{2}} v_{pq} w_{pq}
  = \v{\m v} \c\v{\m w},
\end{align}
and simplify
\begin{align}
  \diagram{two_body_1}
  &= \diagram{two_body_1_o} - \diagram{two_body_1_x} \\
  &= \diagram{two_body_1_oo} - \diagram{two_body_1_ox}
  - \diagram{two_body_2} \label{eq:mid_zero} \\
  &= \v v \c \v w - 2 \v{\m v} \c \v{\m w},
\end{align}
where the middle diagram in \eqref{eq:mid_zero} contains an empty sum
for the cross, and is therefore equal to zero.  Similarly neglecting
diagrams with empty sums and additionally using the fact that
$v_{pq},w_{pq}$ are mean-zero, we can work out
\begin{align}
  \diagram{two_body_0}
  &= \diagram{two_body_0_o} - \diagram{two_body_0_x} \\
  &= \sp{\diagram{two_body_0_oo} - \diagram{two_body_0_ox}}
  - 2 \sp{\diagram{two_body_1}} \\
  &= \sp{-2 \diagram{two_body_1_o}}
  - 2 \sp{\diagram{two_body_1_oo} - \diagram{two_body_2}} \\
  &= -4 \diagram{two_body_1_oo} + 2 \diagram{two_body_2} \\
  &= -4 \v v\c\v w + 4 \v{\m v}\c\v{\m w}.
\end{align}
Altogether, we thus have that
\begin{multline}
  \sum_{k<\ell} \sum_{p<q} v_{k\ell} w_{pq} O_{k\ell} Q_{pq} \\
  \EQFS - \p{\v v \c \v w - \v{\m v} \c \v{\m w}} O_{1,2} Q_{3,4}
  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} O_{1,2} Q_{1,3}
  + \v{\m v} \c \v{\m w}\, O_{1,2} Q_{1,2}.
  \label{eq:POQP}
\end{multline}

@@

Here we simplify a product of three two-body operators of a particular
form projected onto the permutationally symmetric manifold:
\begin{align}
  \X \equiv \P_0 \sp{\sum_{k<\ell} \sum_{p<q} \sum_{r<s}
    u_{k\ell} v_{pq} w_{rs} X_k X_\ell X_p X_q X_r X_s} \P_0,
  &&
  X^2 = 1,
\end{align}
where $k,\ell,p,q,r,s$ index individual spins; $X$ is a single-spin
operator; and $u_{k\ell},v_{pq},w_{rs}$ are arbitrary scalar
coefficients with $m_{pq}=m_{qp}$ and $m_{pp}=0$ for each of
$m\in\set{u,v,w}$.  Collecting terms according to the number of $X$
operators that remain after accounting for $X^2=1$, we find that
\begin{align}
  \X \EQFS A_6 X^{\otimes 6} + A_4 X^{\otimes 4} + A_2 X^{\otimes 2},
  \label{eq:triple_multi}
\end{align}
with
\begin{align}
  A_6 \equiv \f18 \diagram{triple_0},
  &&
  A_4 \equiv \diagram{triple_1} + \f12 \diagram{triple_01},
  \label{eq:triple_64}
\end{align}
\begin{align}
  A_2 \equiv \f12 \diagram{triple_2} + \diagram{triple_11}
  + \f14 \diagram{triple_02} + \diagram{triple_011},
  \label{eq:triple_2}
\end{align}
where we have neglected scalar terms without physical consequence, and
a diagram with unlabeled circles denotes a sum over all distinct label
assignments, e.g.
\begin{align}
  \diagram{triple_1} \equiv \diagram{triple_1_uvw},
  \label{eq:triple_uvw}
\end{align}
\begin{align}
  \diagram{triple_011}
  \equiv \diagram{triple_011_uvw}
  + \diagram{triple_011_vwu} + \diagram{triple_011_wuv}.
  \label{eq:triple_uvw_3}
\end{align}
The first diagram in each of $A_6,A_4,A_2$ as written in
\eqref{eq:triple_64}, \eqref{eq:triple_2} thus has only one assignment
of labels, as in \eqref{eq:triple_uvw}, while all other unlabeled
diagrams have three assignments, as in \eqref{eq:triple_uvw_3}.

>>>>

To simplify the product in \eqref{eq:sym_prod_proj}, we first write
\begin{align}
  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
  = \prod_{j\in\ZZ_p} \f1{M_j!} \sum_{k\in\ZZ_N^{M_j}}
  w_j\p{k} O_j\p{k},
  \label{eq:sym_prod_vec}
\end{align}
where we extend the definitions of $w_j\p{k}$ and $O_j\p{k}$ to all
vectors $k\in\ZZ_N^{M_j}$ by defining symmetric dimension-$M_j$
tensors $w_j$ and $O_j$ (i.e.~tensors with $M_j$ indices, with $O_j$
an {\it operator-valued} tensor) for which $w_j\p{k}=O_j\p{k}=0$ if
$k$ contains any repeated symbols.  We then collect terms whose
operator content is equivalent up to a permutation of spins.  To this
end, we classify terms in \eqref{eq:sym_prod_vec} by the numbers $g_S$
of indices shared by all tensors $w_j,O_j$ with $j\in S\subset\ZZ_p$.
For example, a term of the form
$w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}$ with distinct indices
$\p{a,b,c,d,e}\in\C_N\p{5}$ would have
\begin{align}
  g_{\set{1}} = 1,
  &&
  g_{\set{1,2,3}} = 1,
  &&
  g_{\set{1,3}} = 1,
  &&
  g_{\set{2,3}} = 2,
\end{align}
and $g_S=0$ for all other subsets $S\subset\ZZ_3$.  This index
assignment can be represented by the Venn diagram
\begin{align}
  \diagram{example_123},
  \label{eq:venn_diagram}
\end{align}
where $g_S$ is determined the number of dots at the intersection of
circles $j\in S$.  We denote an assignment of distinct indices
according to choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ by $g$, where
$\PP\p{\ZZ_p}$ is the power set (i.e.~set of all subsets) of $\ZZ_p$.
Given a vector $\m M$ of the dimensions $M_j$, we then denote the set
of all valid index assignments $g$ by $\G\p{\m M}$.  For consistency,
all valid index assignments $g\in\G\p{\m M}$ must assign exactly $M_j$
indices to the dimension-$M_j$ tensor $w_j$, i.e.~
\begin{align}
  \sum_{S\in\PP\p{\ZZ_p}\,:\,j\in S} g_S
  = \sum_{R\in\PP\p{\ZZ_p\setminus\set{j}}} g_{\set{j}\cup R}
  = M_j
\end{align}
for all $j\in\ZZ_p$.  In order to make the index assignments
determined by choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ unique, we
also enforce that $g_{\set{}}=0$ for all $g\in\G\p{\m M}$, which is
equivalent to saying that every dot in the corresponding Venn diagram
must lie within some circle.  The set of valid index assignments
$\G\p{\m M}$ is then essentially the set of all Venn diagrams of the
form in \eqref{eq:venn_diagram} with $p=\dim\p{\m M}$ circles and a
total of $M_j$ dots within circle $j$.

@@

The simplest diagrams have $g_S$ dots (``$\bullet$'') at the
intersection of the circles $j\in S$, and are simply equal to the
associated scalar $w\p{g}$ defined in \eqref{eq:diagram_factors}, e.g.
\begin{align}
  \diagram{example_123}
  \equiv \sum_{\substack{a,b,c,\in\ZZ_N\\\p{d,e}\in\C_N\p{2}\\
      \abs{\set{a,b,c,d,e}}=5}}
  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}.
\end{align}
Such a diagram with $\abs{g}$ dots nominally takes $\O\p{N^{\abs{g}}}$
time to compute.  In practice, this computational cost can be
prohibitive for performing numerical simulations of systems with any
appreciable size.  We therefore wish to simplify diagrams in order to
reduce their computational cost as much as possible.  Our general
strategy will be to replace constrained sums that make all indices
interdependent (such as $\abs{\set{a,b,c,d,e}}=5$ above) by
unconstrained sums that can be carried out in sequence.  In order to
represent different constraints, we need to define diagrams in which
filled dots (``$\bullet$'') may be replaced by empty dots
(``$\circ$'') or crosses (``$\bm\times$'').

If a region of a diagram contains $F$ filled dots, those dots
correspond to indices that sum over $\C_N\p{F}$ with the constraint
that they are not equal to any other filled dots in a diagram.  $E$
empty dots in a region, meanwhile, correspond to indices that are
summed over $\C_N\p{E}$ without any additional constraints, e.g.
\begin{align}
  \diagram{example_o}
  \equiv \sum_{\substack{\p{a,b}\in\C_N\p{2}\\e\in\ZZ_N\\
      \abs{\set{a,b,e}}=3}}
  \sum_{\substack{\p{c,d}\in\C_N\p{2}\\f\in\ZZ_N}}
  v\p{a,b,c,d,e} w\p{e,f}.
\end{align}
Empty dots are therefore entirely independent of filled dots.
Crosses, meanwhile, are {\it entirely dependent} on filled dots:
``cross indices'' are summed {\it only} over the values of ``filled
dot indices'', e.g.
\begin{align}
  \diagram{example_x}
  \equiv \sum_{\substack{a,e\in\ZZ_N\\\abs{\set{a,e}}=2}}
  \sum_{b,c\in\set{a,e}} \sum_{d\in\ZZ_N}
  u\p{a,b,c} v\p{b,d,e} w\p{b,c,e}.
\end{align}
Empty dots and crosses allow us to decompose diagrams with a high
computational cost into different diagrams with a lower computational
cost.  We will essentially use two tricks to decompose diagrams:
eliminating filled dots, and eliminating crosses.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Decomposing diagrams}

Given a diagram containing only dots (filled or empty), we can
decompose any filled dot into an empty dot and a cross.  For example
\begin{align}
  \diagram{example_elim}
  = \diagram{example_elim_o}
  - \diagram{example_elim_x}
  \label{eq:example_elim}
\end{align}
where
\begin{align}
  \diagram{example_elim}
  \equiv \sum_{\substack{a,b,c\in\ZZ_N\\\abs{\set{a,b,c}}=2}}
  v\p{a,b} w\p{b,c},
\end{align}
\begin{align}
  \diagram{example_elim_o}
  \equiv \sum_{\substack{b,c\in\ZZ_N\\\abs{\set{b,c}}=2}}
  v\p{\circ,b} w\p{b,c},
  &&
  v\p{\circ,b} \equiv \sum_{a\in\ZZ_N} v\p{a,b},
  \label{eq:example_elim_o}
\end{align}
\begin{align}
  \diagram{example_elim_x}
  \equiv \sum_{\substack{b,c\in\ZZ_N\\\abs{\set{b,c}}=2}}
  \sum_{a\in\set{b,c}} v\p{a,b} w\p{b,c}.
\end{align}
The decomposition in \eqref{eq:example_elim} essentially breaks
up a sum over $a\in\ZZ_N\setminus\set{b,c}$ into a difference of sums
over $a\in\ZZ_N$ and $a\in\set{b,c}$.  The sum over $a\in\ZZ_N$ to
compute $v\p{\circ,b}$ has $\O\p{N^2}$ cost, and can be performed
prior to evaluating the diagram in \eqref{eq:example_elim_o}.
The decomposition in \eqref{eq:example_elim} therefore splits
one $\O\p{N^3}$ into two $\O\p{N^2}$ diagrams.

After splitting a filled dot into an empty dot and a cross, the next
step is to immediately eliminate the cross.  To eliminate a cross from
a diagram, we first note that the corresponding index can ignore other
indices on tensors that contain the index; that is (pay attention to
the sum over $a$),
\begin{align}
  \diagram{example_elim_x}
  \equiv \sum_{\substack{b,c\in\ZZ_N\\\abs{\set{b,c}}=2}}
  \sum_{a\in\set{b,c}} v\p{a,b} w\p{b,c}
  = \sum_{\substack{b,c\in\ZZ_N\\\abs{\set{b,c}}=2}}
  \sum_{a\in\set{c}} v\p{a,b} w\p{b,c}.
\end{align}
For each remaining index addressed by a cross, meanwhile, the cross
simply forces the corresponding indices to be equal, which is
equivalent to moving a filled dot into a different region:
\begin{align}
  \diagram{example_elim_x}
  = 2 \diagram{example_elim_x_full}
  = 2 \sum_{b,c\in\C_N\p{2}} v\p{c,b} w\p{b,c}.
  \label{eq:example_elim_final}
\end{align}
The factor of 2 accounts for the fact that the left diagram sums over
$b,c\in\ZZ_N$, while the right diagram sums over
$\p{b,c}\in\C_N\p{2}$.  We can generally repeat the above process of
eliminating filled dots and crosses until only empty dots remain in a
diagram.  The only complication in carrying out this process is that
we have to manually keep track of symmetry factors at each step, such
as the factor of 2 in \eqref{eq:example_elim_final}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Symmetry factors}

Here we discuss the rules governing symmetry factors that appear when
decomposing diagrams.  In general, eliminating a filled dot from a
region with $F$ filled dots makes a diagram pick up a factor of $1/F$,
e.g.
\begin{align}
  \diagram{example_sym}
  = \f14 \diagram{example_sym_o}
  - \f14 \diagram{example_sym_x}.
\end{align}
This factor can be derived by replacing the sum over $\C_N\p{F}$ for
the filled dots within the region by a sum over $\ZZ_N^{F}$, which
picks up a factor of $1/F!$ to account for the number of ways to
permute $F$ values.  Setting aside one filled dot for replacement by
an empty dot and a cross, the sum over $\ZZ_N^{F-1}$ for the remaining
dots can be changed back to a sum over $\C_N\p{F-1}$, picking up a
factor of $\p{F-1}!$.  The overall factor picked up throughout this
procedure is therefore $\p{F-1}!/F!=1/F$.  Similarly, adding an empty
dot to a region with $E$ empty dots makes the diagram pick up a factor
of $\p{E+1}!/E!=E+1$, so
\begin{align}
  \diagram{example_sym_o}
  = \f23 \diagram{example_sym_oo}
  - \f13 \diagram{example_sym_ox}.
\end{align}
Finally, eliminating a cross by moving a filled dot from an ``old''
region with $F_{\t{old}}$ filled dots into a ``new'' region with
$F_{\t{new}}$ filled dots picks up an overall factor of $F_{\t{new}}$,
so
\begin{align}
  \diagram{example_sym_x}
  = 2 \diagram{example_sym_x_elim},
\end{align}
where the factor of $1/F_{\t{old}}$ that would be acquired from
removing a dot from the old region is canceled out by a factor of
$F_{\t{old}}$ that accounts for the number of ways to move a dot from
the old region to the new one.

@@

Calculating the second-order effective Hamiltonian $H_2^{(2)}$ induced
on the permutationally symmetric manifold $\M_0$ by the perturbation
$\V_2$ in \eqref{eq:perturbations} requires us to simplify a product
of the form
\begin{multline}
  \P_0 \sp{\sum_{k<\ell} \sum_{p<q}
    v_{k\ell} w_{pq} O_{k\ell} Q_{pq}} \P_0 \\
  \EQFS \diagram{two_body_0} O_{1,2} Q_{3,4}
  + \diagram{two_body_1} O_{1,2} Q_{1,3}
  + \diagram{two_body_2} O_{1,2} Q_{1,2},
  \label{eq:POQP_start}
\end{multline}
where $k,\ell,p,q$ index individual spins; $O$ and $Q$ are
permutationally-symmetric two-spin operators; the coefficients
$v_{k\ell},w_{pq}$ satisfy \eqref{eq:cond_2} with eigenvalue
$\Delta\ne0$, which implies that they are mean-zero; and the factors
of $1/4$ and $1/2$ are included to cancel out the symmetry factors
that appear in the definition of a diagram.  Defining
\begin{align}
  \v{\m m} \equiv \sum_{p<q} m_{pq} \ket{\set{p,q}},
  &&
  \v m \equiv \sum_p m_p \ket{p},
  &&
  m_p \equiv \sum_q m_{pq},
\end{align}
for $m\in\set{v,w}$, we can write
\begin{align}
  \diagram{two_body_2} \equiv \sum_{p,q\in\C_N\p{2}} v_{pq} w_{pq}
  = \v{\m v} \c\v{\m w},
\end{align}
and simplify
\begin{align}
  \diagram{two_body_1}
  &= \sp{\diagram{two_body_1_o}} - \sp{\diagram{two_body_1_x}} \\
  &= \sp{\diagram{two_body_1_oo} - \diagram{two_body_1_ox}}
  - \sp{2\diagram{two_body_2}} \label{eq:mid_zero} \\
  &= \v v \c \v w - 2 \v{\m v} \c \v{\m w},
\end{align}
where the middle diagram in \eqref{eq:mid_zero} contains an empty sum
for the cross, and is therefore equal to zero.  Similarly neglecting
diagrams with empty sums and additionally using the fact that
$v_{pq},w_{pq}$ are mean-zero, we can work out
\begin{align}
  \diagram{two_body_0}
  &= \f12 \sp{\diagram{two_body_0_o}}
  - \f12 \sp{\diagram{two_body_0_x}} \\
  &= \f12 \sp{\diagram{two_body_0_oo} - \diagram{two_body_0_ox}}
  - \f12 \sp{\diagram{two_body_1}} \\
  &= \f12 \sp{-\diagram{two_body_1_o}}
  - \f12 \p{\v v\c\v w - 2\v{\m v}\c\v{\m w}} \\
  &= -\v v\c\v w + \v{\m v}\c\v{\m w}.
\end{align}
Altogether, we thus have that
\begin{multline}
  \sum_{k<\ell} \sum_{p<q} v_{k\ell} w_{pq} O_{k\ell} Q_{pq} \\
  \EQFS - \p{\v v \c \v w - \v{\m v} \c \v{\m w}} O_{1,2} Q_{3,4}
  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} O_{1,2} Q_{1,3}
  + \v{\m v} \c \v{\m w}\, O_{1,2} Q_{1,2}.
  \label{eq:POQP}
\end{multline}

@@

Here we simplify a product of three two-body operators of a particular
form projected onto the permutationally symmetric manifold:
\begin{align}
  \X \equiv \P_0 \sp{\sum_{k<\ell} \sum_{p<q} \sum_{r<s}
    u_{k\ell} v_{pq} w_{rs} X_k X_\ell X_p X_q X_r X_s} \P_0,
  &&
  X^2 = 1,
\end{align}
where $k,\ell,p,q,r,s$ index individual spins; $X$ is a single-spin
operator; and $u_{k\ell},v_{pq},w_{rs}$ are arbitrary scalar
coefficients with $m_{pq}=m_{qp}$ and $m_{pp}=0$ for each of
$m\in\set{u,v,w}$.  Collecting terms according to the number of $X$
operators that remain after accounting for $X^2=1$, we find that
\begin{align}
  \X \EQFS A_6 X^{\otimes 6} + A_4 X^{\otimes 4} + A_2 X^{\otimes 2},
  \label{eq:triple_multi}
\end{align}
with
\begin{align}
  A_6 \equiv \diagram{triple_0},
  &&
  A_4 \equiv \diagram{triple_01} + \diagram{triple_1},
  \label{eq:triple_64}
\end{align}
\begin{align}
  A_2 \equiv \diagram{triple_011} + \diagram{triple_02}
  + \diagram{triple_11} + \diagram{triple_2},
  \label{eq:triple_2}
\end{align}
where we have neglected scalar terms without physical consequence, and
a diagram with unlabeled circles denotes a sum over all distinct label
assignments, e.g.
\begin{align}
  \diagram{triple_1} \equiv \diagram{triple_1_uvw},
  \label{eq:triple_uvw}
\end{align}
\begin{align}
  \diagram{triple_011}
  \equiv \diagram{triple_011_uvw}
  + \diagram{triple_011_vwu} + \diagram{triple_011_wuv}.
  \label{eq:triple_uvw_3}
\end{align}
The first diagram in each of $A_6,A_4,A_2$ as written in
\eqref{eq:triple_64}, \eqref{eq:triple_2} thus has only one assignment
of labels, as in \eqref{eq:triple_uvw}, while all other unlabeled
diagrams have three assignments, as in \eqref{eq:triple_uvw_3}.

Having identified diagrammatic representations of the coefficients
$A_6,A_4,A_2$, we now need to simplify all diagrams to reduce their
computational complexity, as discussed in Section \ref{sec:diagrams}.
Simplifying all these diagrams by hand is tedious, but the process of
eliminating filled dots and crosses from a diagram is algorithmic
enough to execute on a computer, doing which we find
\begin{align}
  A_6 = D_0 - D_1 + D_2 - D_3,
  &&
  A_4 = D_1 - 2 D_2 + 3 D_3,
  &&
  A_2 = D_2 - 3 D_3,
\end{align}
where
\begin{align}
  D_0 \equiv \diagram{triple_0_o},
  &&
  D_1 \equiv \diagram{triple_01_o} - 2 \diagram{triple_1_o},
  &&
  D_3 \equiv \diagram{triple_0111_o},
\end{align}
\begin{align}
  D_2 \equiv \diagram{triple_011_o}
  + \diagram{triple_02_o}
  - 2 \diagram{triple_11_o}
  + 4 \diagram{triple_2_o}.
\end{align}
Defining, for each of $m\in\set{u,v,w}$,
\begin{align}
  \m m \equiv \sum_{p,q} m_{pq} \op{p}{q},
  &&
  \v{\m m} \equiv \sum_{p<q} m_{pq} \ket{\set{p,q}},
  &&
  \v m \equiv \sum_p m_p \ket{p},
  &&
  m_p \equiv \sum_q m_{pq},
  &&
  \col{m} \equiv \sum_{p<q} m_{pq},
\end{align}
we can expand
\begin{align}
  D_0 = \col{u}\,\col{v}\,\col{w},
  &&
  D_1 = \col{u}\,\v v\c\v w + \col{v}\,\v w\c\v u
  + \col{w}\,\v u\c\v v - 2 \sum_p u_p v_p w_p,
  &&
  D_3 = \tr\p{\m u \c \m v \c \m w},
\end{align}
\begin{multline}
  D_2 = \v u \c\m v\c\v w + \v v \c\m w\c\v u + \v w \c\m u\c\v v
  + \col{u}\,\v{\m v}\c\v{\m w} + \col{v}\,\v{\m w}\c\v{\m u}
  + \col{w}\,\v{\m u}\c\v{\m v} \\
  - 2 \sum_{p,q} \p{u_p v_{pq} w_{pq}
    + v_p w_{pq} u_{pq} + w_p u_{pq} v_{pq}}
  + 4 \sum_{p<q} u_{pq} v_{pq} w_{pq}.
\end{multline}
Finally, we can also write the product $\X$ in \eqref{eq:triple_multi}
in terms of the collective operator $\col{X} \equiv \sum_p X_p$:
\begin{align}
  \X \EQFS
  \tilde A_6 \col{X}^6 + \tilde A_4 \col{X}^4 + \tilde A_2 \col{X}^2,
  \label{eq:triple_col}
\end{align}
where we have neglected scalar terms without physical consequence, and
\begin{align}
  \tilde A_6 \equiv \f{A_6}{N!_6},
  &&
  \tilde A_4 \equiv \f{A_4}{N!_4} - 5\p{3N-8} \f{A_6}{N!_6},
\end{align}
\begin{align}
  \tilde A_2 \equiv \f{A_2}{N!_2} - 2\p{3N-4} \f{A_4}{N!_4}
  + \p{45N^2-210N+184} \f{A_6}{N!_6}.
\end{align}
